### 2021-03-15 report/submission.tex
Report updated. Conclusions and references added. Intro remaining. Optionally, discussion of fourier transforms can also be added.

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in, footskip=0.25in]{geometry}

\usepackage{graphicx}
\graphicspath{../}

\usepackage{booktabs}
\usepackage{wrapfig}
\usepackage{subcaption}

\usepackage[backend=bibtex, style=numeric, sorting=none]{biblatex}
\bibliography{references}

\usepackage[font=footnotesize,labelfont=bf]{caption}
\captionsetup{width=\columnwidth}

\usepackage{amsmath}

\begin{document}

	% ======================= start titlepage ==============================================

	\begin{titlepage}
		\begin{center}
			\Large Numerical Methods in Astrophysics \\
			\vspace{1cm}
			\huge{
				Project 2 \\
				\vspace{0.5cm}
				\textbf{Ordinary Differential Equations,}\\
				\textbf{Fourier Transforms and} \\
				\textbf{Spectral Analysis} \\
				\vspace{1cm}
			}
			\Large \emph{Saksham Kaushal}
		\end{center}
	\end{titlepage}

	% =========================end titlepage ================================================

	\tableofcontents
	\newpage

	% ========================= Introduction ================================================

	\section{Introduction} \label{Introduction}

	This section introduced the principles of numerical methods involved in this study. In sections \ref{task1} and \ref{task2} the methods, results and discussion of observations are presented for part 1 and 2 of the project respectively. Section \ref{conclusions} gives a brief summary of physical phenomena studies using these simulations along with conclusions derived from both parts of the project.

	% ========================= Part 1 ======================================================

	\newpage
	\section{Part 1 - Ordinary Differential Equations} \label{task1}

	For stars in disk of the galaxy, equation of motion is given by equation 19 of \cite{lecture_3} as,
	\begin{equation}
		\frac{d^2r(t)}{dt^2} = - \nabla \Phi (r) -2 \Omega_B \times v+ \Omega_B^2 r ,
		\label{eq:motion1}
	\end{equation}

	where \(\Omega_B\) is angular velocity and the gravitational potential, \(\Phi (r)\) in the two dimensions for an elongated bar structure is given by equation 20 of \cite{lecture_3} as,
	\begin{equation}
		\Phi(x,y)=\frac{v_0^2}{2} ln(a^2+x^2+\frac{y^2}{q^2}). 
		\label{eq:grav_potential}
	\end{equation}

	The gradients of gravitational potential along the two axes are given by partial derivatives of equation (\ref{eq:grav_potential}) with respect to \(x\) and \(y\) as follows,
	\begin{equation}
		\frac{\partial \Phi}{\partial x} = \frac{v_0^2 x}{a^2+x^2+\frac{y^2}{q^2}},
		\label{eq:grad_grav_potential_x}
	\end{equation}
	\begin{equation}
		\frac{\partial \Phi}{\partial y} = \frac{v_0^2 y}{q^2(a^2+x^2+\frac{y^2}{q^2})}.
		\label{eq:grad_grav_potential_y}
	\end{equation}

	We can substitute equations (\ref{eq:grad_grav_potential_x}) and (\ref{eq:grad_grav_potential_y}) in equation (\ref{eq:motion1}) to obtain the equation of motion that can be solved in our simulation. Equations along the two axes are given by,
	\begin{equation}
		a_x = \frac{d^2x}{dt^2} = - \frac{v_0^2 x}{a^2+x^2+\frac{y^2}{q^2}} + 2\Omega_B v_y + x \Omega_B^2 ,
		\label{eq:acceleration_x}
	\end{equation}
	\begin{equation}
		a_y = \frac{d^2y}{dt^2} = - \frac{v_0^2 y}{q^2(a^2+x^2+\frac{y^2}{q^2})} - 2\Omega_B v_x + y \Omega_B^2,
		\label{eq:acceleration_y}
	\end{equation}

	where it must be noted that for second term on the right hand side of equation (\ref{eq:acceleration_x}), the cross product \(\Omega_B \times v_x = -v_y\) and for similar term in equation (\ref{eq:acceleration_y}), \(\Omega_B \times v_y = v_x\), where \(\Omega_B\) has the direction perpendicularly outwards from the x-y plane.

	The value of \emph{Jacobi integral} is conserved as a star moves in a rotating barred potential and  therefore, minimum dispersion in its values throughout the simulation indicates better numerical integration. Its value is given by 
	\begin{equation}
		E_j \equiv \frac{\dot r^2}{2} + \Phi -\frac{(\Omega_B \times r)^2}{2}.
		\label{eq:jacobi_integral}
	\end{equation}

	From equation 13 of \cite{lecture_3}, the numerical error in each integration time-step can be controlled by adapting the step-size accordingly. If we define,
	\begin{equation}
		\Delta_0 \equiv \epsilon y_{max},
		\label{eq:numerical_error}
	\end{equation}
	then \(\Delta < \Delta_0\) should be satisfied for each integration step, where \(\Delta\) is the numerical error of a given step, \(\epsilon\) is the parameter defining the accuracy of integration and \(y_{max}\) is the amplitude of function in the range over which function is being integrated. If this inequality does not hold true, the values need to be adjusted accordingly, and the integration step is repeated with reduced step-size until this inequality is satisfied. However, if this inequality does indeed hold true such that \(\Delta \ll \Delta_0\), integration step-size can be safely increased to reduce the difference between \(\Delta\) and \(\Delta_0\). The step-size can be update according to equation 14 of \cite{lecture_3}, given as,
	\begin{equation}
		h = h_n \left| \frac{\Delta_0}{\Delta} \right| ^m,
		\label{eq:stepsize}
	\end{equation}
	where \(h\) is the new updated step-size, \(h_n\) is the previous step-size and \(m\) is an exponent usually set to different values when increasing or decreasing the step-size, often in the range \([0.2,0.25]\).

		% ----------------------------Methods -----------------------------------------------

		\subsection{Methods} \label{methods1}

		In the FORTRAN code \emph{orbit19.f}, the four elements of array \texttt{ys} represent the values of \(x\), \(v_x\), \(y\) and \(v_y\) respectively. Value of \(a\), \(q\), \(v_0\) and \(\Omega_B\) are represented by variables \texttt{a, q, v0} and \texttt{om}. The difference between the variables \texttt{x2} and \texttt{x1} indicates the time for which the simulation takes place. This is essentially different from time-step of the simulation, that depends on the value of \(\epsilon\), which parameterizes the accuracy of integration. The four elements of array \texttt{dxdy} represent the values of \(v_x\), \(a_x\), \(v_y\) and \(a_y\). The values of \(a_x\) and \(a_y\) are calculated using the equations (\ref{eq:acceleration_x}) and (\ref{eq:acceleration_y}), and encoded in the subroutine \texttt{derivs} in the code. The Jacobi integral, given by equation (\ref{eq:jacobi_integral}) is assigned the variable \texttt{ej}. \\

		The following methods were employed for different tasks performed for this part of the project:
		\begin{enumerate}

			\item The missing \textit{x} and \textit{y} components of the equation of motion for a star in barred potential well are encoded in the subroutine \texttt{derivs} of FORTRAN program \textit{orbit19.f} as, \\

			\noindent \texttt{dydx(2) = -y(1)*v0**2/(a**2+y(1)**2+(y(3)**2/q**2))+2*om*y(4)+y(1)*om**2}, and \\
			\texttt{dydx(4) = -y(3)*v0**2/(q**2*(a**2+y(1)**2)+y(3)**2)-2*om*y(2)+y(3)*om**2} . \\

			These are the \(a_x\) and \(a_y\) components, given by equations (\ref{eq:acceleration_x}) and (\ref{eq:acceleration_y}). The \(v_x\) and \(v_y\) components were already assigned to array elements \texttt{dxdy(1)} and \texttt{dxdy(3)} respectively.

			\item Initial parameters of the simulation for this task, along with their values and variable names in the program are mentioned in Table \ref{table:task1.2}.

			\begin{table} 
				\centering
				\begin{tabular} {c c c}
					\toprule
					\textbf{Parameter} & \textbf{Variable} & \textbf{Value} \\
					\midrule
					\(\Omega_B\) (Angular velocity) & \texttt{om} & 1.0 \\
					\(v_0\) & \texttt{v0} & 1.0 \\
					\(a\) & \texttt{a} & 0.03 \\
					\(q\) & \texttt{q} & 0.8 \\
					\(t_1\) (Initial time) & \texttt{x1} & 0.0 \\
					\(t_2\) (Final time) & \texttt{x2} & 5.0 \\
					\(x\) & \texttt{ys(1)} & 0.0 \\
					\(v_x\) & \texttt{ys(2)} & -1.68 \\
					\(y\) & \texttt{ys(3)} & 0.087 \\
					\(v_y\) & \texttt{ys(4)} & 0.0 \\
					\bottomrule
				\end{tabular}
			\caption{Initial parameters of the simulation, variable names assigned to them in the FORTRAN code and their initial values set for the first simulation.}
			\label{table:task1.2}
			\end{table}

			\item The value of \(\epsilon\) varies the number of time-steps involved, and therefore controls the accuracy of integration. For the initial conditions given in Table \ref{table:task1.2}, simulation is run for values of \(\epsilon\) equal to \(10^{-5}, 10^{-6}, 10^{-7}, 10^{-8}\) and \(10^{-9}\). The value of Jacobi integral, defined above, is inspected for each of these values of \(\epsilon\) and plotted on a graph using Python and its associated libraries like numpy, pandas, matplotlib and seaborn. All subsequent plotting is performed in a similar manner.

			\item The value of variable \texttt{x2}, defining the final time of simulation, is changed from initial value of \(5.0\) to a new value of \(20.0\), keeping all other parameters unaltered. The trajectories calculated in both the cases are then plotted.

			\item The value of \(v_x\) assigned to array element \texttt{ys(2)} is changed to values -1.64, -1.635, -1.63, -1.58, -1.15, -0.6, -0.524 and -0.4, and the simulation is run for each of these values to obtain the corresponding trajectories which are plotted for a final time of 20.0.

		\end{enumerate}

		% ----------------------------Results -----------------------------------------------

		\subsection{Results} \label{results1}

		\begin{itemize}

			\item For different values of \(\epsilon\), the variations in values of Jacobi integral are plotted in figure \ref{fig:task1.3} upto a final time of 5.0. The variations are most prominent for highest and lowest values of \(\epsilon\), and least for intermediate values. Analytically, the Jacobi integral is a conserved quantity for a three-body problem in a non-inertial frame of reference. Therefore, minimum variation in its value corresponds to most accurate numerical integration, which eventually gives us the most suitable value of \(\epsilon\).

			From the graph in figure \ref{fig:task1.3} and using the values of standard deviations for different values of \(\epsilon\) from table \ref{table:task1.3}, we can conclude that \(\epsilon = 10^{-7}\) is the most suitable value for integration.

			\begin{figure} [h]
				\includegraphics[width=\columnwidth]{../task1/q3-epsilon/plots/q3.jpg}
				\caption{Variation of Jacobi integral with time for different values of \(\epsilon\), a parameter that signifies the accuracy of integration.}
				\label{fig:task1.3}
			\end{figure}

			\begin{table} [h]
				\centering
				\setlength{\tabcolsep}{20pt} 
				\begin{tabular} {c c}
					\toprule
					\textbf{\(\epsilon\)} & \textbf{\(\sigma_{E_j}\)} \\
					\midrule
					\addlinespace
					\(10^{-5}\) & \(7.326 \times 10^{-6}\)
					\\ 
					\(10^{-6}\) & \(6.490 \times 10^{-7}\) \\
					\(10^{-7}\) & \(1.887 \times 10^{-7}\) \\
					\(10^{-8}\) & \(6.918 \times 10^{-7}\) \\
					\(10^{-9}\) & \(1.013 \times 10^{-5}\) \\
					\bottomrule
				\end{tabular}
				\caption{Standard deviation \((\sigma_{E_j})\) in values of Jacobi Integral for different values of integration accuracy parameter \((\epsilon)\). }
				\label{table:task1.3}
			\end{table}

			\item The trajectories for a star with initial parameters defined in section \ref{methods1}, for final times of 5.0 and 20.0 in simulation are plotted in figure \ref{fig:task1.4}.

			\begin{figure} 
				\includegraphics[width=\columnwidth]{../task1/q4-trajectory/plots/q4.jpg}
				\caption{Trajectories followed over (a) 5.0 units and (b) 20 units of time by a star with initial parameters, \((x,y)=(0,0.087)\) and \((v_x,v_y)=(-1.68,0)\).}
				\label{fig:task1.4}
			\end{figure}

			\item The calculated trajectories for a final time of 20.0 corresponding to initial values of \(v_x = \) -1.68, -1.64 and -1.635 are plotted for comparison in figure \ref{fig:task1.5}-(Top). The individual trajectories are plotted separately in figures \ref{fig:task1.4}(b) and \ref{fig:task1.5}-(Bottom).

			\begin{figure} [h]
				\includegraphics[width=\columnwidth]{../task1/q5q6-velocity/plots/q5_combined.jpg}
				\hrule
				\includegraphics[width=\columnwidth]{../task1/q5q6-velocity/plots/q5_sep.jpg}
				\caption{(Top) A comparison plot showing trajectories followed by a star for different initial values of \(v_x\). (Bottom \(1 \times 2\) grid) Trajectories followed by a star for initial values of \(v_x\) equal to (a) -1.64, and (b) -1.635 shown separately. Note that the trajectory for \(v_x = -1.68\) is shown individually in figure \ref{fig:task1.4}(b). }
				\label{fig:task1.5}
			\end{figure}

			\item Similar plots for initial values of \(v_x = \) -1.63, -1.58, -1.15, -0.6, -0.524 and -0.4 are plotted for comparison in figure \ref{fig:task1.6}-(Top) and separately in figure \ref{fig:task1.6}-(Bottom).

			\begin{figure} [h]
				\includegraphics[width=\columnwidth]{../task1/q5q6-velocity/plots/q6_combined.jpg}
				\hrule
				\includegraphics[width=\columnwidth]{../task1/q5q6-velocity/plots/q6_sep.jpg}
				\caption{(Top) A comparison plot showing trajectories followed by a star for different initial values of \(v_x\). (Bottom \(3 \times 2\) grid) Trajectories followed by a star for initial values of \(v_x\) equal to (a) -1.63, (b) -1.58, (c) -1-15, (d) -0.6, (e) -0.524 and (f) -0.4 shown separately.}
				\label{fig:task1.6}
			\end{figure}
		\end{itemize}

		% ----------------------------Discussions -----------------------------------------------

		\subsection{Discussions} \label{discussions1}

		\begin{enumerate}

			\item The accuracy of integration increases with reduction in time-step, although it increases execution time. However, below a certain value of integration accuracy parameter, \(\epsilon\), the accuracy tends to decrease, as seen in figure \ref{fig:task1.3}.

			\item The trajectory followed by the star with \(v_x = -1.635\) in figure \ref{fig:task1.5}(b) is a case of \emph{stable periodic orbit} \cite{orbits_lamfr}. Such an orbit has a closed geometry and the motion repeats after a fixed interval of time.

			\item Slight variation in the parameters from the case of \(v_x = -1.635\), we get \emph{trapped orbits} \cite{orbits_lamfr}. Examples of such orbits are for \(v_x = -1.64\) and \(-1.68\). A larger deviation in parameters from the stable periodic orbit leads to a less tightly bound trapped orbit, as in the case of \(v_x = -1.68\). Such orbits are neither periodic nor closed and tend to stay in the vicinity of the parent stable periodic orbit.

			\item Similar trend can be noted in figure \ref{fig:task1.6}, where the stable periodic orbit occurs at \(v_x = -0.524\) (figure \ref{fig:task1.6}(e)). A morphological transition in orbits can be seen as we move from one stable periodic orbit (\(v_x = -1.635\)) to another (\(v_x = -0.524\)). Accordingly, the orbit for \(v_x = -1.15\), for which the initial parameters deviate the most, from the two periodic stable orbits, ends up being the trapped orbit with most loosely bound morphology.

			\item With decrease in \(v_x\), we get orbits that are more tightly bound to the center of the potential well. This trend can be seen through the decreasing span of orbits along x-axis as we move to lower values of \(v_x\) in figures \ref{fig:task1.5} and \ref{fig:task1.6}.

		\end{enumerate}

	% ========================= Part 1 ======================================================

	\clearpage
	\section{Part 2 - Fourier Transform and Spectral Analysis} \label{task2}

		% ----------------------------Methods -----------------------------------------------

		\subsection{Methods} \label{methods2}

		In the FORTRAN code \emph{fourier19.f}, the setup for the simulation is performed in a manner similar to that in section \ref{methods1}, with \(v_x = -1.635\). The simulation produces output files containing values of the frequency running number (\(k\)), the power spectral density as percentage (\(prw\)) and the logarithm of this value (\(\log{prw}\)). The value of \(k\) can be converted to frequency, \(\omega\) using the relation,
		\begin{equation}
			\omega = \frac{2\pi (k-1)}{KMAXX*step},
			\label{eq:k_to_freq}
		\end{equation} 
		where \(KMAXX = 2^{13} = 8192\) and \(step = 0.02\) are predefined in the program.

		The value of \(v_x\), assigned to variable \texttt{ys(2)}, is set to \(-1.635\), \(-1.64\), \(-1.68\) and \(-1.15\) in separate runs of the simulation, each producing a different output file to be analyzed.

		% ----------------------------Results -----------------------------------------------

		\subsection{Results} \label{results2}

		\begin{figure}
			\includegraphics[width=\columnwidth]{../task2/plots/q1-5_lin.jpg}
			\caption{One-sided power spectral density (in \%) vs frequency plots for various initial values of \(v_x\).}
			\label{fig:task2_lin}
		\end{figure}

		\begin{table}

			\begin{subtable} [h] {0.275\linewidth}
				\centering
				\begin{tabular} {l l l}
					\toprule
					\textbf{k} & \textbf{\(\omega\)} &  \\
					\midrule
					123 & 4.678 & \(a\) \\
					245 & 9.357 & \(\approx 2a\) \\
					367 & 14.036 & \(\approx 3a\) \\
					488 & 18.676 & \(\approx 4a\) \\
					610 & 23.355 & \(\approx 5a\) \\
					732 & 28.03 & \(\approx 6a\) \\
					854 & 32.71 & \(\approx 7a\) \\
					\(\dots\) & & \\
					\bottomrule
				\end{tabular}
				\caption{\(v_x = -1.635\).}
				\label{table:task2_1}
			\end{subtable}%
			\hspace{0.05\linewidth}
			\begin{subtable} [h] {0.275\linewidth}
				\centering
				\begin{tabular} {l l l}
					\toprule
					\textbf{k} & \textbf{\(\omega\)} &  \\
					\midrule
					98 & 3.72 & \(b\) \\
					121 & 4.602 & \(a\) \\
					144 & 5.484 & \(\approx 2a-b\) \\
					218 & 8.322 & \(\approx a+b\) \\
					240 & 9.165 &  \(\approx 2a\)\\
					264 & 10.086 & \(\approx 3a-b\) \\
					337 & 12.885 & \(\approx 2a+b\) \\
					360 & 13.767 & \(\approx 3a\) \\
					383 & 14.65 & \(\approx 4a-b\) \\
					457 & 17.526 & \(\approx 3a+b\) \\
					480 & 18.37 & \(\approx 4a\) \\
					503 & 19.29 & \(\approx 5a-b\) \\
					\(\dots\) & & \\
					\bottomrule
				\end{tabular}
				\caption{\(v_x = -1.64\).}
				\label{table:task2_2}
			\end{subtable}
			\hspace{0.05\linewidth}
			\begin{subtable} [h] {0.275\linewidth}
				\centering
				\begin{tabular} {l l l l}
					\toprule
					& \textbf{k} & \textbf{\(\omega\)} &  \\
					\midrule
					* & 20 & 0.73 & \(b\) \\
					& 64 & 2.416 & \(\approx a-2b\) \\
					* & 83 & 3.145 & \(\approx a-b\) \\
					* & 102 & 3.873 & \(a\) \\
					* & 122 & 4.64 & \(\approx a+b\) \\
					* & 141 & 5.37 & \(\approx a+2b\) \\
					& 166 & 6.328 & \(\approx 2a-2b\) \\
					* & 185 & 7.056 & \(\approx 2a-b\) \\
					* & 204 & 7.785 & \(\approx 2a\) \\
					* & 223 & 8.514 & \(\approx 2a+b\) \\
					* & 242 & 9.242 & \(\approx 2a+2b\) \\
					& 267 & 10.2 & \(\approx 3a-2b\) \\          
					& 281 & 10.738 & \(\approx 3a-b\) \\ 
					& 305 & 11.66 & \(\approx 3a\) \\
					* & 324 & 12.387 & \(\approx 3a+b\) \\
					& 344 & 13.154 & \(\approx 3a+2b\) \\
					& \(\dots\) &  & \\
					\bottomrule
				\end{tabular}
				\caption{\(v_x = -1.68\). The starred rows comprise of peaks with a component \(>0.1\%\) of total power.}
				\label{table:task2_3}
			\end{subtable}
			\caption{Frequency running number (\(k\)) and corresponding frequency values (\(\omega\)) for few initial peaks in the data for different initial values of \(v_x\). Algebraic representation of accompanying dependent frequencies in terms of independent frequencies is also given in the unlabeled rightmost column.}
			\label{table:task2_1-3}
		\end{table}

		\begin{figure}
			\includegraphics[width=\columnwidth]{../task2/plots/q1-5_log.jpg}
			\caption{Logarithm (base 10) of one-sided power spectral density vs frequency plots for various initial values of \(v_x\).}
			\label{fig:task2_log}
		\end{figure}

		\begin{table}
			\centering
			\begin{tabular} {l l l}
				\toprule
				\textbf{k} & \textbf{\(\omega\)} & \textbf{\% of total power} \\
				\midrule
				289 & 11.045 & 35.177 \\ 
				360 & 13.767 & 3.23 \\
				647 & 24.77 & 1.73 \\
				356 & 13.61 & 1.214 \\
				\bottomrule
			\end{tabular}
			\caption{Frequencies and contribution to total power of major peaks in power spectrum for \(v_x=-1.15\).}
			\label{table:task2_5}
		\end{table}

		\begin{itemize}

			\item One-sided PSD for \(v_x = -1.635\) is plotted on linear scale in figure \ref{fig:task2_lin}(a) and on logarithmic scale in figure \ref{fig:task2_log}(a). The initial few major peaks are listed in table \ref{table:task2_1}.

			\item One-sided PSD for \(v_x = -1.64\) is plotted on linear scale in figure \ref{fig:task2_lin}(b) and on logarithmic scale in figure \ref{fig:task2_log}(b). The initial few major peaks are listed in table \ref{table:task2_2}.

			\item One-sided PSD for \(v_x = -1.68\) is plotted on linear scale in figure \ref{fig:task2_lin}(c) and on logarithmic scale in figure \ref{fig:task2_log}(c). The initial few major peaks are listed in table \ref{table:task2_3}.

			\item One-sided PSD for \(v_x = -1.15\) is plotted on linear scale in figure \ref{fig:task2_lin}(d) and on logarithmic scale in figure \ref{fig:task2_log}(d). The frequencies of major peaks along with the values of how much percentage of total power they constitute are given in table \ref{table:task2_5}.

		\end{itemize}

		% ----------------------------Discussions -----------------------------------------------

		\subsection{Discussions} \label{discussions2}

		\begin{enumerate}

			\item \begin{enumerate}
				\item Given that there is almost no power present in high frequencies, no significant aliasing is expected in this case.
				\item There is no power at zero frequency and all power peaks lie at frequencies significantly higher than Nyquist frequency (\(= \frac{2 \pi}{N \Delta}\), where \(N=\)\texttt{KMAXX} and \(\Delta=\)\texttt{step} in the code). This indicates that unresolved peaks in power at low frequencies are not expected.
				\item As can be seen from figure \ref{fig:task2_log}(a), the distribution of power spectral density consists of a set of sharp peaks.
				\item From figure \ref{fig:task2_log}(a), it can be observed that the peaks do not occur randomly and their distribution across the range of frequencies seems regular. This can be confirmed from table \ref{table:task2_1}, where we see that these peaks occur at regular intervals of a frequency. 
			\end{enumerate}

			\item Table \ref{table:task2_1} enlists the frequencies at which peaks in power occur. Relations of these peaks to the primary peak are given in the last column, from which we can infer that peaks occur at frequencies which are integral multiples of frequency of the primary peak. No secondary peak is required to explain any peak in the sequence. Also, from the data, it is observed that the power in these peaks decreases with increase in the integral coefficient, in accordance with the property of Fourier transforms given in \cite{lecture_4}. The frequency, \(\omega\) of the primary peak is 4.678 (\(k=123\)). Therefore all subsequent peaks can be represented as, \(\omega_n=4.678n\), where \(n=2, 3, 4, \dots\). Hence the trajectory oscillates with a single frequency, \(\omega = 4.678\).

			From section \ref{discussions1}, we know that in this case, the star is in a stable periodic orbit. From preceding discussion we also know that this trajectory oscillates with a single frequency. Thus, we expect the star to be at the same point after each complete period of oscillation. Hence, after one period of oscillation, the star sis located at the same point from where it started. This is confirmed by finding the coordinates of star after \(t=\frac{2\pi}{\omega}\). Given \(\omega=4.678\), we get \(t \approx 1.343\). By interpolation of data obtained in section \ref{results1} for \(v_x=-1.625\), we obtain \((x,y)={-0.0021,-0.0870}\) at \(t=1.343\). For comparison, \((x,y)=(0.0000,0..870)\) at \(t=0\) almost coincides with this value. Therefore, the star is observed at same location after one period.

			\item \begin{enumerate}
				\item Similar to result 1 above, no significant aliasing or unresolved peaks at low frequencies are expected in this case also, for reasons mentioned in part 1(a) and 1(b) respectively. 
				\item There exists some pattern in frequencies of the power peaks, however, as seen from table \ref{table:task2_2}, the pattern is much more complicated than for peaks in table \ref{table:task2_1}. As such, the pattern cannot be explained using a single frequency.
				\item From the last column in table \ref{table:task2_2}, we can see that all major peaks can be explained using additional peaks due two main frequencies, i.e. 4.602 (\(k=121\)) and 3.72 (\(k=98\)). 
				\item The power in these two frequencies is \(\approx 71.384\%\) and \(\approx 0.093\%\) of total power respectively. However, some of the power of these peaks is distributed in nearby frequencies also.
			\end{enumerate}

			\item \begin{enumerate}
				\item For figure \ref{fig:task2_log}(c), there exist several well-defined peaks, but the pattern in those peaks is almost incomprehensible. The number of those peaks is also much larger than in case of figure \ref{fig:task2_log}(b). As a consequence, percentage contribution of these peaks towards total power is also less in this case.
				\item Each peak containing more than \(0.1\%\) of total power is marked with an asterisk in table \ref{table:task2_3}. All these peaks can be explained using a linear combination of two major peaks at frequencies 3.873 and 0.73, as can be seen from last column of table \ref{table:task2_3}.
				\item These two independent frequencies contain \(\approx 48.022\%\) and \(\approx 0.325\%\) of total power.   
			\end{enumerate}

		\end{enumerate}

	% ========================= Conclusions =================================================

	\section{Conclusions} \label{conclusions}

	In this project we study the orbits of objects in barred potential wells, which represent the physical scenario of stars in the bars of the galaxies. The trajectories vary substantially as we alter parameters like initial position and velocity. On the quantitative analysis side of our study, how precisely we calculate these trajectories has a profound impact on our observations and understandings.

	The ordinary differential equations can be numerically integrated using several methods. In this project, equation of motion of star in a potential was integrated using fourth order Range-Kutta method. The accuracy and precision of our integration was tested by analyzing the values of a quantity conserved over time, known as the Jacobi integral. Orbits for various combinations of parameters were obtained and their morphology was examined.

	A subset of these orbits was then analyzed using Fourier transforms. From power spectrum analysis of these orbits, we obtained the distribution of power in orbits across a range of frequencies.

	We conclude that stable periodic orbits oscillate at a single frequency and therefore a star in such an orbit is expected to undergo periodic motion over a closed trajectory. On the other hand, trapped orbits oscillate at a combination of multiple frequencies. A star in such an orbit tends to remain in vicinity of its parent stable periodic orbit and follows a trajectory with open geometry. The more the initial parameters mentioned above vary for a star in trapped orbit, the lesser it tends to remain in vicinity of parent stable periodic orbit. Also, with decrease in velocity, a star follows an orbit that spans less farther away from the center of potential well.

	% ========================= References =================================================

	\printbibliography[heading=bibintoc]

\end{document}